GLOBALIZER: A NOVEL SUPERCOMPUTER SOFTWARE SYSTEM FOR SOLVING TIME-CONSUMING GLOBAL OPTIMIZATION PROBLEMS

### 2016-08-16 2016/GOM/paper.tex
Fix authors list in GOM paper
Some minor corrections

\author{
\name{V. P. Gergel\textsuperscript{a}$^{\ast}$\thanks{$^\ast$Corresponding author. Email: gergel@unn.ru}
and K. A. Barkalov\textsuperscript{b} and A. V. Sysoev}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK;
\textsuperscript{b}State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
}

@@

  D=\{y\in \mathbf{R}^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}

@@

corresponding minimized function to satisfy the uniform Hölder condition (see \cite{strongin1978}, \cite{strSergGO}) i.e.

@@

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
		\caption{A Peano curve approximation for the third level of}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensional reduction consists of the following (see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x)\),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

@@

f(x) = \varphi(y(x)):x\in [0,1]\}

@@

0=x_0<x_1<\dotsc<x_{k+1}=1

@@

the points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,

@@

\textit{Rule 2}. To compute a current estimate of the Hölder constant \(H\) from (\ref{eq:holder})

@@

  R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1 \\
  R(i)=2\rho_i-4\dfrac{z_i}{M},\quad i=1 \\
  R(i)=2\rho_{i}-4\dfrac{z_{i-1}}{M},\quad i=k+1

@@

R(t)=\max_{1\leqslant i \leqslant k+1}R(i)

@@

  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{M}\right]^N,\quad 1<t<k+1 \\
  x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1

@@

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leqslant j\leqslant p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leqslant j\leqslant p\), located in the
intervals with the highest characteristics from (\ref{step4par})
\begin{equation}
\label{step5par}
\begin{array}{cr}\\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1 \\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1
\end{array}
\end{equation}
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in \cite{strSergGO}.
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

@@

  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s

>>>>

\author{
\name{V.P. Gergel and K.A. Barkalov$^{\ast}$\thanks{$^\ast$Corresponding author. Email: barkalov@vmk.unn.ru} and A.V. Sysoyev}
\affil{State University of Nizhny Novgorod, Nizhny Novgorod, Russia}
}

@@

  D=\{y\in \mathbf{R}^N:a_i\leqslant y_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}

@@

corresponding minimized function to satisfy the uniform H{\"o}lder condition (see \cite{strongin1978}, \cite{strSergGO}) i.e.

@@

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
		\caption{A Peano curve approximation for the third level}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensional reduction consists of the following (see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional function \(\varphi(y(x))\),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) in the mapping \(y(x)\) is calculated,
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the multidimensional point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

@@

f(x) = \varphi(y(x)):x\in [0,1].

@@

0=x_0<x_1<\dotsc<x_{k+1}=1.

@@

The points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,

@@

\textit{Rule 2}. To compute a current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})

@@

  R(i)=\rho_i+\dfrac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\dfrac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
  R(i)=2\rho_i-4\dfrac{z_i}{M},\quad i=1, \\
  R(i)=2\rho_{i}-4\dfrac{z_{i-1}}{M},\quad i=k+1.

@@

R(t)=\max_{1\leqslant i \leqslant k+1}R(i).

@@

  x_{k+1}=\dfrac{x_t+x_{t-1}}{2}-\sign(z_{t}-z_{t-1})\dfrac{1}{2r}\left[\dfrac{|z_{t}-z_{t-1}|}{M}\right]^N,\quad 1<t<k+1, \\
  x_{k+1}=\dfrac{x_t+x_{t-1}}{2},\quad t=k+1.

@@

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leqslant i\leqslant k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geqslant R(t_2)\geqslant \dots \geqslant R(t_{k})\geqslant R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leqslant j\leqslant p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leqslant j\leqslant p\), located in the
intervals with the highest characteristics from (\ref{step4par})
\begin{equation}
\label{step5par}
\begin{array}{cr}\\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2}-\sign(z_{t_j}-z_{t_j-1})\dfrac{1}{2r}\left[\dfrac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\quad 1<t_j<k+1, \\
x_{k+j}=\dfrac{x_{t_j}+x_{t_j-1}}{2},\quad t_j=1,t_j=k+1.
\end{array}
\end{equation}
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leqslant j\leqslant p.
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in \cite{strSergGO}.
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

@@

  \varphi(y^l(x^*))=\min\{\varphi(y^l(x)):x\in [0,1]\},1\leqslant l\leqslant s.
